Galaxy Zoo: Morphological Classifications for Galaxies in HST Legacy Imaging

### 2016-04-27 writeup/gz_hubble_data.tex
COSMOS comparisons - Tasca and Cassata

One of the major difficulties in studying the morphologies of galaxies in large samples lies in the system of measurement. Visual classification by experts has been used for many years \citep[eg,][]{hub26,dev59,san61,van76,nai10,bai11,kar15}. These methods have advantages in using the significant processing power of the human brain to identify patterns, but suffer from issues such as lack of scaling to large surveys and potential issues with replicability and calibration. Automated measurements, both parametric \citep{pen02a,sim11,lac12} and non-parametric \citep{con03,lot04,sca07,bam08,fre13}, scale well to arbitrary sample sizes, but do not always fully capture the relevant features, especially for asymmetric galaxies that become increasingly common at high redshifts. The Galaxy Zoo project \citep{lin08,lin11} utilizes crowdsourced visual classifications to measure galaxies in color-composite images. With $>2\times10^5$~classifiers, this allows for multiple independent classifications of each image which are combined and calibrated to give a distribution of vote fractions proportional to the probability of a feature being visible. While the crowdsourced data require extensive calibration \citep{bam09,wil13}, they have a proven reliability and have been used in dozens of papers \citep[eg,][]{lan08,bam09,dar10,mas11c,ski12,sim13,sch14,wil15}.

@@

Since the majority of the data in Legacy surveys described in \S\ref{ssec:legacy_surveys} had images in only 1 or 2~bands available when GZH was launched, making standard 3-channel RGB images in the same method used for the SDSS images in the original Galaxy Zoo was not possible. For surveys with images in two bands (AEGIS, GEMS, and the two-epoch GOODS-N and GOODS-S), the lower-wavelength band is mapped to the blue channel, the higher-wavelength band to the red channel, and the green channel created by taking the geometric mean of the two. Bands used in each of the surveys are listed in Table~\ref{tbl:gzh_numbers}. Note that the 2-epoch GOODS-N and GOODS-S images use different filters --- this was a deliberate choice made so that the GEMS images could be directly compared with the overlapping coverage of GOODS-S (Figure~\ref{fig:filtercurves}).

@@

\begin{figure*}
\includegraphics[width=160mm]{figures/gzh_interface.png}
\caption{Screenshot of the Galaxy Zoo: Hubble interface showing an example COSMOS image at the first step in the decision tree.\label{fig:interface}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/gzh_decision_tree.pdf}
\caption{Flowchart of the questions presented to GZH users, labeled with the corresponding Task numbers. Tasks in the decision tree are color-coded by tier level: Gray-colored Tasks are 1st tier questions which are asked of all users. Tasks colored green, blue, purple, and pink (respectively) are one, two, three, or four steps below branching points in the decision tree.}
\label{fig:decisiontree}
\end{center}
\end{figure*}

@@

\begin{figure*}
\includegraphics[width=160mm]{figures/example_ferengi.pdf}
\caption{Examples of two galaxies which have been run through the FERENGI code to produce simulated \hst{} images. The value of \pfeatures{} for each panel is (1) Top row: \pfeatures~= 0.9, 0.625, 0.35, 0.35, 0.225 and (2) Bottom row: \pfeatures~= 1.00, 0.875, 0.875, 0.625, 0.375. \label{fig:exampleFERENGI}}
\end{figure*}

@@

The intended purpose of the GZH catalog is to provide a simple, yet flexible, means of identifying samples of galaxies with a desired morphological type. Here we provide instructions for creating such samples using the vote fractions corresponding to the tasks shown in Figure~\ref{fig:decisiontree}. We stress that the selection process will vary based on the particular science case. More conservative cuts can be applied to create pure, but not necessarily complete, samples. These are useful for selecting galaxies exhibiting unique morphologies for individual case studies or observing proposals. Looser cuts can be applied to obtain samples with a higher level of completeness, which will increase the sample size but can potentially decrease purity. Population studies would make use of such large samples, in which statistical significance is a crucial factor in evaluating results. \citet{mas11c,mel14}, and \citet{gal15} are examples of GZ papers which select morphologies in this way.

In GZH, volunteers answer questions about a galaxy's morphology in a decision tree format. With this structure, questions shown to a user are dependent on their answers to the previous questions. For example, a user is only asked Task 11, ``How many spiral arms?'' if they had answered ``yes'' to Task 10, which asks whether any spiral arms are present. This structure is used to reduce the workload of the users by only asking questions relevant to the particular galaxy; in this example, it would not be useful, nor would it make physical sense, to ask a user how many spiral arms in a galaxy which they had declared had no spiral arms to begin with. Figure~\ref{fig:decisiontree} shows the possible paths available for answering the questions offered in GZH. The colors represent the tier level of Tasks, which indicate the number of previous Tasks the given Task is dependent upon. The arm-number question, Task 11, is a 5th tier Task, meaning that whether this question is seen by a user is dependent on four Tasks preceding it. In this case, it is only shown to volunteers who voted the galaxy was featured/disk-shaped in Task 01, not clumpy in Task 12, not edge-on in Task 02, and had spiral structure in Task 10.

To select galaxies of a morphological type identified with a particular Task, a cut is placed on the vote fraction for that Task ($p_{\rm task}$), \emph{as well as} the vote fractions for the Tasks preceding it, because of the dependency induced by the decision-tree structure. For example, to select barred galaxies, a cut may be placed on $p_{\rm bar}$ such that only galaxies where a high fraction of votes for this task voted for the $bar,yes$ answer. This is not the only necessary cut, however, since not all users answer this question; only those who have previously selected ``features'' in Task~01, ``not clumpy'' in Task~12, and ``not edge-on'' in Task~02 will have the opportunity to vote on the bar question, Task~03. To ensure that $p_{\rm bar}$ is well-sampled, cuts on all previous tasks must be applied.

The flexibility of this catalog allows the user to set their own selection criteria for vote fraction thresholds to create a morphologically-pure sample. In Table~\ref{tbl:thresholds} we offer baseline cuts for selecting galaxies of a variety of morphologies. We detemined these thresholds by a visual inspection of subsamples. For each Task, we first required that at least 20 users voted on the given question. We then applied a cut on the vote fraction for the previous task and analyzed a subsample of 50 galaxies meeting these criteria, as well as a control sample of galaxies which had 20 users vote on the task, but did not meet the threshold cut set for the previous task. The threshold cut was adjusted and new subsamples were inspected until both the original and control samples achieved $\ge 80\%$ purity.

As an example of how to use Table~\ref{tbl:thresholds} to create a sample of 3-armed spirals galaxies, one should select objects with $N_{arm~number} \ge 20$, $p_{\rm features}>0.23, p_{\rm clumpy,no} > 0.30, p_{edgeon,no}>0.25$, and $p_{spiral,yes}>0.25$. These cuts define a sample of galaxies of ``arm number candidates''; ie galaxies for which answering the arm number question makes physical sense and the vote fraction $p_{\rm arm~number}$ is well-sampled. In this example, such galaxies are featured, non-clumpy, non-edge~on, spiral galaxies. At this point a cut can be made on $p_{\rm arm~number=3}$ to select spirals with three arms.

Tasks 03, 04 and 05 have an additional possible pathway; as shown in Figure~\ref{fig:decisiontree}, a user might also be shown this question if they select ``featured/disk'' in Task 01, ``clumpy'' in Task 12, two or more clumps in Task 16, and ``spiral arrangement'' in Task 15. By applying the appropriate thresholds for this path, we find that $< 0.5\%$ of the galaxies which have $\ge 20$ answers to these questions used this pathway to arrive at these Tasks. Further, of these subjects, none were found to exhibit disk structure, although the clumps within were arranged in a spiral pattern. We therefore do not find this pathway necessary in examining responses to Tasks 03, 04, and 05.

In this section we have described how to use the previous Task thresholds in Table~\ref{tbl:thresholds} to select well-sampled galaxies which may or may not contain the feature associated with a unique Task. We will now offer two examples of how to use the vote fractions for the final Task to obtain a sample of galaxies with a certain morpholgical type, by creating a sample of bars and samples of clumpy galaxies grouped by clump mulitiplicity.

\subsection{Example 1: Selecting bars}

We created a sample of barred galaxies by applying cuts on the previous tasks as listed in Table~\ref{tbl:thresholds}. 11,049 ``bar candidates'', galaxies for which asking the bar question is meaningful, were selected by applying the cuts $N_{bar} \ge 20$, $p_{\rm features}>0.23, p_{\rm clumpy,no} > 0.30,$ and $p_{edgeon,no}>0.25$. These galaxies are featured, non-clumpy, non-edge~on galaxies. Of these, a pure sample of 730 barred disks was identified by applying a cut of $p_{bar}>0.7$. A subsample of 50 galaxies were visually inspected and 94\% were found to contain strong bars. A complete sample of strong and weak bars was created by applying a cut of $p_{bar}>0.3$. This sample contained 3,218 galaxies, 86\% of which were found to contain weak or strong bars through visual inspection in a subsample of 50.

We use the complete bar sample to estimate the redshift evolution of bar fraction and find a steady decrease of $f_{bar} \sim 0.32$ at $z=0.4$ to $f_{bar} \sim 0.24$ at $z=1.0$. The decrease in bar fraction agrees with \ref{mel14}, however they report a lower bar fraction overall, finding a decrease from $f_{bar}=0.22$ at $z=0.4$ to $f_{bar}=0.11$ at $z=1.0$. The difference in total bar fraction is expected, as our analysis used a looser cut on $p_{\rm bar}$, we did not apply a luminosity cut, and we use debiased values for $p_{\rm features}$ which increases the total amount of disks in the sample. For a more thorough analysis of the bar fraction, a cut on $p_{\rm bar}$ that evolves with redshift may yield more accurate results, since there are no explicit debiased values of $p_{\rm bar}$ that would take redshift induced bias into account.

\subsection{Example 2: Identifying clump multiplicity}
For galaxies identified as ``clumpy'' in GZH, the number of clumps can be determined using Task 16. We identify samples of galaxies having one, two, three, four, and more than four clumps. Using Table~\ref{tbl:thresholds}, we identify 8,444 clumpy galaxies using $p_{\rm features} > 0.23$ and $p_{\rm clumpy,yes}>0.80$ to ensure the vote fractions for Task 16 are well-sampled. We were able to identify with reasonable confidence the clump number of 1,112 of the clumpy galaxies; in the remainder, the unique clumps were less distinguishable from each other and the exact number of clumps could not be deduced without careful visual inspection. In the 1,112 which did have distinguishable clumps, we identified 61 one-clump galaxies using $p_{\rm 1~clump}>0.50$ 442 two-clump galaxies using $p_{\rm 2~clumps}>0.80$, 275 three-clump galaxies using $p_{\rm 3~clumps}>0.75$, 71 four-clump galaxies using $p_{\rm 4~clumps} > 0.70$, and 263 galaxies with more than four clumps using $p_{\rm >4~clumps}>0.70$. Alternatively, these data may be used to create more general samples of clumpy galaxies with few clumps and many clumps. We define a sample of 989 ``few clumps'' galaxies using $(p_{\rm 1~clump} + p_{\rm 2~clumps}) > 0.5$ and 2,910 ``many clumps'' galaxies using $(p_{\rm 3~clumps}+p_{\rm 4~clumps}+p_{\rm >4~clumps})>0.5$.

{\note
\begin{itemize}
    \item \sout{explain decision tree format; vote fractions are dependent on previous task vote fractions} 
    \item \sout{provide example of how to select bars}
    \item \sout{discuss difference between conservative/clean sample and loose/complete sample, emphasize goals differ depending on science case}
    \item \sout{science with bar example - estimate bar fraction vs redshift}
    \item science with clump multiplicity - not sure actually 
\end{itemize}
} 
\begin{table}
\caption{Suggested thresholds for selecting morphological samples from Galaxy Zoo: Hubble \label{tbl:thresholds}}
\begin{tabular}{llll}
\hline\hline
No.      &  Task 	            & Previous              & Vote fraction threshold            \\
         &            	        & task(s)               & $N_{\rm task} \ge 20$              \\
\hline
01       & smooth or features   & $-$                   & $-$                                \\
02       & edge on              & 01, 12                & $p_{\rm clumpy,no}>0.30$            \\
03       & bar		            & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
04       & spiral arms          & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
05       & bulge prominence     & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
06       & odd yes/no           & $-$                   & $-$                                \\
07       & rounded              & 01                    & $p_{\rm smooth}>0.70$               \\
08       & odd feature          & 06                    & $p_{\rm odd,yes}> 0.50$            \\
09       & bulge shape          & 01, 12, 02            & $p_{\rm edgeon,yes}>0.40$           \\
10       & arms winding         & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
11       & arms number          & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
12       & clumpy               & 01                    & $p_{\rm features}>0.23$            \\
13       & bright clump         & 01, 12, 16            & $p_{\rm one~clump}< 0.40$           \\
14       & bright central clump & 01, 12, 16, 13        & $p_{\rm bright~clump,yes}>0.50$     \\
15       & clump arrangement    & 01, 12, 16            & $p_{\rm multiple~clumps}>0.45$     \\
16       & clump count          & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
17       & clumps symmetrical   & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
18       & clumps embedded      & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
\hline\hline
\end{tabular}
\end{table}

@@

\subsection{Comparison to other catalogs}\ref{ssec:comparisons}

@@

>>>>

One of the major difficulties in studying the morphologies of galaxies in large samples lies in the system of measurement. Visual classification by experts has been used for many years \citep[eg,][]{hub26,dev59,san61,van76,nai10,bai11,kar15}. These methods have advantages in using the significant processing power of the human brain to identify patterns, but suffer from issues such as lack of scaling to large surveys and potential issues with replicability and calibration. Automated measurements, both parametric \citep{pen02a,sim11,lac12} and non-parametric \citep{abr03,con03,lot04,sca07,bam08,fre13}, scale well to arbitrary sample sizes, but do not always fully capture the relevant features, especially for asymmetric galaxies that become increasingly common at high redshifts. The Galaxy Zoo project \citep{lin08,lin11} utilizes crowdsourced visual classifications to measure galaxies in color-composite images. With $>2\times10^5$~classifiers, this allows for multiple independent classifications of each image which are combined and calibrated to give a distribution of vote fractions proportional to the probability of a feature being visible. While the crowdsourced data require extensive calibration \citep{bam09,wil13}, they have a proven reliability and have been used in dozens of papers \citep[eg,][]{lan08,bam09,dar10,mas11c,ski12,sim13,sch14,wil15}.

@@

Since the majority of the data in Legacy surveys described in \S\ref{ssec:legacy_surveys} had images in only 1 or 2~bands available when GZH was launched, making standard 3-channel RGB images in the same method used for the SDSS images in the original Galaxy Zoo was not possible. For surveys with images in two bands (AEGIS, GEMS, and the two-epoch GOODS-N and GOODS-S), the lower-wavelength band is mapped to the blue channel, the higher-wavelength band to the red channel, and the green channel created by taking the geometric mean of the two. Bands used in each of the surveys are listed in Table~\ref{tbl:gzh_numbers}. The 2-epoch GOODS-N and GOODS-S images use different filters --- this was a deliberate choice made so that the GEMS images could be directly compared with the overlapping coverage of GOODS-S (Figure~\ref{fig:filtercurves}).

@@

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/gzh_interface.png}
\caption{Screenshot of the Galaxy Zoo: Hubble interface showing an example COSMOS image at the first step in the decision tree.\label{fig:interface}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/gzh_decision_tree.pdf}
\caption{}
\label{fig:decisiontree}
\end{center}
\end{figure*}

@@

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/example_ferengi.pdf}
\caption{Examples of two galaxies which have been run through the FERENGI code to produce simulated \hst{} images. The value of \pfeatures{} for each panel is (1) Top row: \pfeatures~= 0.9, 0.625, 0.35, 0.35, 0.225 and (2) Bottom row: \pfeatures~= 1.00, 0.875, 0.875, 0.625, 0.375. \label{fig:exampleFERENGI}}
\end{figure*}

@@

The intended purpose of the GZH catalog is to provide a simple, yet flexible, means of identifying samples of galaxies with a desired morphological type. Here we provide instructions for creating such samples using the vote fractions corresponding to the tasks shown in Figure~\ref{fig:decisiontree}. We stress that the selection process will vary based on the particular science case.

To select galaxies of a morphological type identified with a Task offered in the decision tree (see Figure~\ref{fig:decisiontree}, a cut is placed on the vote fraction for that Task ($p_{\rm task}$), as well as the vote fractions for the Tasks preceding it. For example, to select barred galaxies, a cut may be placed on $p_{\rm bar}$ such that only galaxies where a high fraction of votes for this task voted for the $bar,yes$ answer. This is not the only necessary cut, however, since not all users answer this question; only those who have previously selected ``features'' in Task~01, ``not clumpy'' in Task~12, and ``not edge-on'', \emph{or} previously selected ``features'' in Task~01, ``clumpy'' in Task~12, ``$>$ 2 clumps'' in Task~16, and ``spiral arrangement'' in Task~15 will have the opportunity to vote on the bar question, Task~03. To ensure that $p_{\rm bar}$ is well-sampled, cuts on all previous tasks must be applied.

The flexibility of this catalog allows the user to set their own selection criteria for vote fraction thresholds to create a morphologically-pure sample. In Table~\ref{tbl:thresholds} we offer baseline cuts for selecting galaxies of a variety of morphologies. {\note As an example: to select barred galaxies, one should select objects with $p_{\rm features}>0.52, p_{\rm clumpy,no} > 0.55,$ and $p_{edgeon,no}>x.xx$ \emph{or} objects with $p_{\rm features}>0.52, p_{\rm clumpy,yes} > x.xx, (p_{\rm clump~count, 3}+p_{\rm clump~count,4}+p_{\rm clump~count,>4}+p_{\rm clump~count,can't~tell})>x.xx,$ and $p_{\rm clump~arrangement,spiral}>x.xx.$} These cuts define a sample of galaxies of ``bar candidates''; ie galaxies for which answering the bar question makes physical sense and the vote fraction $p_{\rm bar}$ is well-sampled. In this example, such galaxies are featured, non-clumpy, non-edge~on galaxies, or featured, clumpy, spiral galaxies. At this point a cut can be made on $p_{\rm bar}$ to select a clean sample of only strong bars $p_{\rm bar}>0.8$, or a complete sample of weak and strong bars $p_{\rm bar}>0.3$, or the value most appropriate to the science case.

{\note
\begin{itemize}
    \item explain decision tree format; vote fractions are dependent on previous task vote fractions
    \item provide example of how to select bars
    \item discuss difference between conservative/clean sample and loose/complete sample, emphasize goals differ depending on science case
\end{itemize}
}

\begin{table}
\caption{Suggested thresholds for selecting morphological samples from Galaxy Zoo: Hubble \label{tbl:thresholds}}
\begin{tabular}{llll}
\hline\hline
No.      &  Task 	            & Previous              & Vote fraction threshold            \\
         &            	        & task(s)               & $N_{\rm task} \ge 20$              \\
\hline
01       & smooth or features   & $-$                   & $-$                                \\
02       & edge on              & 01, 12                & $p_{\rm clumpy,no}>0.30$            \\
03       & bar		            & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
04       & spiral arms          & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
05       & bulge prominence     & 01, 12, 02            & $p_{\rm edgeon,no}>0.25$           \\
         &                      & 01, 12, 16, 15        & $p_{\rm clumpy~spiral}>0.65$  \\
06       & odd yes/no           & $-$                   & $-$                                \\
07       & rounded              & 01                    & $p_{\rm smooth}>0.70$               \\
08       & odd feature          & 06                    & $p_{\rm odd,yes}> 0.50$            \\
09       & bulge shape          & 01, 12, 02            & $p_{\rm edgeon,yes}>0.40$           \\
10       & arms winding         & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
11       & arms number          & 01, 12, 02, 04        & $p_{\rm spiral,yes}>0.25$          \\
12       & clumpy               & 01                    & $p_{\rm features}>0.23$            \\
13       & bright clump         & 01, 12, 16            & $p_{\rm one~clump}< 0.40$           \\
14       & bright central clump & 01, 12, 16, 13        & $p_{\rm bright~clump,yes}>0.50$     \\
15       & clump arrangement    & 01, 12, 16            & $p_{\rm multiple~clumps}>0.45$     \\
16       & clump count          & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
17       & clumps symmetrical   & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
18       & clumps embedded      & 01, 12                & $p_{\rm clumpy,yes}>0.80$           \\
\hline\hline
\end{tabular}
\end{table}

@@

\subsection{Comparison to other catalogs}\label{ssec:comparisons}

@@

\subsubsection{COSMOS}

The COSMOS galaxies have several published datasets automatically classifying their morphologies, all using a variation of non-parametric measurements clustered with machine algorithms and trained on a visually-classified reference set of several hundred galaxies. \citet{cas07} use a combination of $C$, $A$, $G$, and $M_{20}$ \citep{cas05} to classify all galaxies with $m_\mathrm{I,petro}<25$. They divide galaxies into three broad morphological categories: early-type, spirals, and irregulars. A very similar method is employed by \citet{tas11}; they use the same non-parametric indices to determine galaxy morphology, but with a slightly different method of calculating the Petrosian radius. 64\% of the galaxies have identical classifications in \citet{cas07} and \citet{tas11}; roughly two-thirds of the disagreements are galaxies originally classified as ``irregular'' in \citet{cas07} but ``spiral'' in \citet{tas11}.

Using a 0.5\arcsec{} positional match, we find 76,241 galaxies with morphologies in both GZH and \citet{cas07}.

% Show the fraction of galaxies in each category as a function of p_features and p_odd?
